---
title: "redo_scientific_data_plots"
output: html_document
date: "2023-03-05"
---

#### This markdown can be used to reproduce the benchmark plots
## Here, we refer each protein matrix dataset as A1, A2, B1 and B2 
## Batch factors and class factors are the same for each cell line pair, so we use metadata from DDA PASEF data for diaPASEF data as well

## A1 = A549/K562 DDA-PASEF 
## A2 = A549/K562 diaPASEF 
## B1 = HCC1806/HS578T DDA PASEF 
## B2 = HCC1806/HS578T diaPASEF


## Libraries
```{r}
library(reshape2)
library(tidyverse)
library(proBatch)
library(sva)
library(visdat)
```

#### Read the A549/K562 datasets 
```{r}
setwd('') #input working directory for stored data

#Read A549/K562 DDA PASEF data 
A1_data <- read.csv('protein_matrix_DDA_A549_K562.csv', header = TRUE, row.names = 1)
#Read A549/K562 diaPASEF data
A2_data <- read.csv('protein_matrix_DIA_A549_K562.csv', header = TRUE, row.names = 1)

#Read metadata 
A1_metadata <- read.csv('metadata_DDA_A549_K562.csv', header = TRUE)
A1_sampleannotation <- read.csv('sampleannotation_DDA_A549_K562.csv', header = TRUE) #metadata for pvca function

#Declare factors 
A1_group <- as.factor(A1_metadata$class) #class factor 
A1_machine <- as.factor(A1_metadata$machine) #batch factor 

#Omit missing value (for use in PCA and PVCA later)
A1_data_nonNA <- na.omit(A1_data) #drop proteins containing missing values in A549/K562 DDA PASEF data 
A2_data_nonNA <- na.omit(A2_data) #drop proteins containing missing values in A549/K562 diaPASEF data
```

#### Read the HCC1806/HS578T datasets 
```{r}
#Read HCC1806/HS578T DDA PASEF data
B1_data <- read.csv('protein_matrix_DDA_HCC1806_HS578T.csv', header = TRUE, row.names = 1)

#Read HCC1806/HS578T diaPASEF data
B2_data <- read.csv('protein_matrix_DIA_HCC1806_HS578T.csv', header = TRUE, row.names = 1)

#Read metadata
B1_metadata <- read.csv('metadata_DDA_HCC1806_HS578T.csv', header = TRUE)
B1_sampleannotation <- read.csv('sampleannotation_DDA_HCC1806_HS578T.csv', header = TRUE) #metadata for pvca function

#Declare factors 
B1_group <- as.factor(B1_metadata$class) #class factor 
B1_machine <- as.factor(B1_metadata$machine) #batch factor 

#Omit missing value (for use in PCA and PVCA later)
B1_data_nonNA <- na.omit(B1_data) #drop proteins containing missing values in HCC1806/HS578T DDA PASEF data 
B2_data_nonNA <- na.omit(B2_data) #drop proteins containing missing values in HCC1806/HS578T diaPASEF data 
```

#Create class specific, batch specfic dataframes for each dataset
```{r}
#A549/K562 datasets 
A549_CAD_A1 <- A1_data[,1:12]
K562_CAD_A1 <- A1_data[,13:24]
A549_N_A1 <- A1_data[,25:36]
K562_N_A1 <- A1_data[,37:48]

A549_CAD_A2 <- A2_data[,1:12]
K562_CAD_A2 <- A2_data[,13:24]
A549_N_A2 <- A2_data[,25:36]
K562_N_A2 <- A2_data[,37:48]

#HCC1806/HS578T datasets 
HCC1806_CAD_B1 <- B1_data[,1:9]
HS578T_CAD_B1 <- B1_data[,10:18]
HCC1806_N_B1 <- B1_data[,19:27]
HS578T_N_B1 <- B1_data[,28:36]

HCC1806_CAD_B2 <- B2_data[,1:9]
HS578T_CAD_B2 <- B2_data[,10:18]
HCC1806_N_B2 <- B2_data[,19:27]
HS578T_N_B2 <- B2_data[,28:36]
```

# Arrange dataframe by class
```{r}
A1_data_ordered <- cbind(A1_data[,1:12], A1_data[25:36], A1_data[,13:24], A1_data[,37:48])
A2_data_ordered <- cbind(A2_data[,1:12], A2_data[25:36], A2_data[,13:24], A2_data[,37:48])
B1_data_ordered <- cbind(B1_data[,1:9], B1_data[19:27], B1_data[,10:18], B1_data[,28:36])
B2_data_ordered <- cbind(B2_data[,1:9], B2_data[19:27], B2_data[,10:18], B2_data[,28:36])
```

## Visualise violin plot representing missing proportions of samples categorised by class and batch 
```{r}
#Function to count missing proportion in each of the samples 
count_na_func <- function(x) sum(is.na(x)/length(x)) 

#Function for violin plot representing missing proportion of samples categorised by class and batch
mv_violin_plot <- function(sample_na_count){ 

NA_violin_plot <- ggplot(melt(sample_na_count), aes(x=variable, y=value, fill = variable)) + 
  geom_violin(size = 0.5) + 
 # geom_boxplot(width = 0.2, size = 1, outlier.size = 0.8) +
  scale_fill_manual(values = c("#469B94", "#469B94", "#F2A83D", "#F2A83D")) + 
  theme_bw() +
  theme(legend.position = "none") + 
  ylab('Missing Value Proportion (%)') +
  xlab('Class (Batch)') + 
  ggtitle('A549/K562 DDA') + 
  theme(plot.title = element_text(size = 21),axis.title.x=element_text(size = 21),axis.title.y=element_text(size = 21)) +
  theme(axis.text.x = element_text(size = 21)) +
  theme(axis.text.y = element_text(size = 21)) +
  theme(axis.title = element_blank()) 
  
return(NA_violin_plot)
}

#### Count NA of datasets 
sample_Na_count_A1 <- round(apply(t(A1_data_ordered), 1, count_na_func), digit = 5) #Count missing proportion in A549/K562 DDA PASEF dataset 
sample_Na_count_A2 <- round(apply(t(A2_data_ordered), 1, count_na_func), digit = 5) #Count missing proportion in A549/K562 diaPASEF dataset 
sample_Na_count_B1 <- round(apply(t(B1_data_ordered), 1, count_na_func), digit = 5) #Count missing proportion in HCC1806/HS578T DDA PASEF dataset 
sample_Na_count_B2 <- round(apply(t(B2_data_ordered), 1, count_na_func), digit = 5) #Count missing proportion in HCC1806/HS578T diaPASEF dataset 

####Categorise by class and batch for each dataset's sample NA count and store in dataframe 
# A549/K562 DDA PASEF dataset 
A1_Na_df_forviol <- data.frame('A549 CAD' = sample_Na_count_A1[1:12], 'A549 N' = sample_Na_count_A1[13:24], 'K562 CAD'= sample_Na_count_A1[25:36], 'K562 N' = sample_Na_count_A1[37:48])

# A549/K562 diaPASEF dataset 
A2_Na_df_forviol <- data.frame('A549 CAD' = sample_Na_count_A2[1:12], 'A549 N' = sample_Na_count_A2[13:24], 'K562 CAD'= sample_Na_count_A2[25:36], 'K562 N' = sample_Na_count_A2[37:48])

# HCC1806/HS578T DDA PASEF dataset 
B1_Na_df_forviol <- data.frame("HCC1806 CAD" = sample_Na_count_B1[1:9], "HCC1806 N" = sample_Na_count_B1[10:18], "HS578T CAD" = sample_Na_count_B1[19:27], "HS578T N" = sample_Na_count_B1[28:36])

# HCC1806/HS578T diaPASEF dataset 
B2_Na_df_forviol <- data.frame("HCC1806 CAD" = sample_Na_count_B2[1:9], "HCC1806 N" = sample_Na_count_B2[10:18], "HS578T CAD" = sample_Na_count_B2[19:27], "HS578T N" = sample_Na_count_B2[28:36])


#### Violin GGPLOT
# A549/K562 DDA PASEF dataset 
mv_violin_plot(A1_Na_df_forviol)

# A549/K562 diaPASEF dataset 
mv_violin_plot(A2_Na_df_forviol)

# HCC1806/HS578T DDA PASEF dataset
mv_violin_plot(B1_Na_df_forviol)

# HCC1806/HS578T diaPASEF dataset 
mv_violin_plot(B2_Na_df_forviol)
```

#Visualise missing value heatmap 
```{r}
## GGPLOT missing values
heatmap_A1 <- vis_miss(A1_data_ordered) + #009BF0", "#C81E27
  scale_fill_manual(values = c('#2C4F8B', '#FD7D79')) +
  ylab('Genes') + 
  theme(axis.text.x = element_text(size = 14, angle = 0)) +  
  theme(axis.text.y = element_text(size = 10)) +
  theme(axis.title = element_text(size = 14)) 

heatmap_A2 <- vis_miss(A2_data_ordered) + #009BF0", "#C81E27
  scale_fill_manual(values = c('#2C4F8B', '#FD7D79')) +
  ylab('Genes') + 
  theme(axis.text.x = element_text(size = 14, angle = 0)) +  
  theme(axis.text.y = element_text(size = 10)) +
  theme(axis.title = element_text(size = 14)) 

heatmap_B1 <- vis_miss(B1_data_ordered) + #009BF0", "#C81E27
  scale_fill_manual(values = c('#2C4F8B', '#FD7D79')) +
  ylab('Genes') + 
  theme(axis.text.x = element_text(size = 14, angle = 0)) +  
  theme(axis.text.y = element_text(size = 10)) +
  theme(axis.title = element_text(size = 14)) 

heatmap_B2 <- vis_miss(B2_data_ordered) + #009BF0", "#C81E27
  scale_fill_manual(values = c('#2C4F8B', '#FD7D79')) +
  ylab('Genes') + 
  theme(axis.text.x = element_text(size = 14, angle = 0)) +  
  theme(axis.text.y = element_text(size = 10)) +
  theme(axis.title = element_text(size = 14)) 
```


#Functions to visualise scatterplots displaying missing rates of proteins ordered by mean log2 intensity
```{r}
#### Function for scatterplot dataframe 
#Obtain mean intensity for each gene
mv_scatter_plot <- function(data){
mean.log2 <-data %>% apply(1, FUN = mean, na.rm = TRUE) %>% c()

#Bind mean intensity to existing df 
data_mean.log2 <-data
data_mean.log2<-cbind(data_mean.log2,mean.log2)

#Sort based on log2intensity | Obtain % missingness | Build scatter df object 
sorted.newdf<-data_mean.log2[order(mean.log2, decreasing =T),]
missingness<-c(rep(0,nrow(data)))
sorted.newdf<-cbind(sorted.newdf,missingness)
for (i in 1:nrow(sorted.newdf)){
  msp<-round((length(which(is.na(sorted.newdf[i,])))/ncol(sorted.newdf))*100,2)
  sorted.newdf[i,'missingness']<-msp
}
scatter.df<-sorted.newdf[,(length(sorted.newdf)-1):length(sorted.newdf)]

#### GGplot function for scatter dataframe
mis_prop_plot <- ggplot(as.data.frame(scatter.df),aes(x=mean.log2,y=missingness))+
  geom_point(size=2,colour='#2C4F8B')+
  labs(title='MV distribution',x='mean log2 protein intensity',y='missing%')+
 # geom_vline(xintercept=median(mean.log2),linetype='longdash')+
  #geom_hline(yintercept=50,linetype='longdash')+
  theme_bw() +
  theme(plot.title = element_text(size = 21),axis.title.x=element_text(size = 21),axis.title.y=element_text(size = 21)) +
  theme(axis.text.x = element_text(size = 21)) +
  theme(axis.text.y = element_text(size = 21)) +
  theme(axis.title = element_blank())

return(mis_prop_plot)
}
```

## Visualise scatterplot displaying missing rates of proteins ordered by mean log2 intensity 
```{r}
## Plot for A549/K562 DDA PASEF 
mv_scatter_plot(A1_data)

## Plot for A549/K562 diaPASEF 
mv_scatter_plot(A2_data)

## Plot for HCC1806/HS578T DDA PASEF 
mv_scatter_plot(B1_data)

## Plot for HCC1806/HS578T diaPASEF 
mv_scatter_plot(B2_data)
```

## Visualise scatterplots demonstrating mean intensities of samples 
```{r}
## Sample mean intensity dataframe for A549/K562 DDA PASEF 
A1_sample_mean_intensity <- data.frame("mean.intensity" = rowMeans(t(na.omit(A1_data)))) #Calculate mean intensity of samples 
A1_sample_mean_df <- data.frame('var' = row.names(A1_sample_mean_intensity), "mean_intensity" = A1_sample_mean_intensity, "class" = factor(A1_group), "batch" = factor(A1_machine))

## Sample mean intensity dataframe for A549/K562 diaPASEF 
A2_sample_mean_intensity <- data.frame("mean.intensity" = rowMeans(t(na.omit(A2_data))))
A2_sample_mean_df <- data.frame('var' = row.names(A2_sample_mean_intensity), "mean_intensity" = A2_sample_mean_intensity, "class" = factor(A1_group), "batch" = factor(A1_machine))

## Sample mean intensity dataframe for HCC1806/HS578T DDA PASEF 
B1_sample_mean_intensity <- data.frame("mean.intensity" = rowMeans(t(na.omit(B1_data))))
B1_sample_mean_df <- data.frame('var' = row.names(B1_sample_mean_intensity), "mean_intensity" = B1_sample_mean_intensity, "class" = factor(B1_group), "batch" = factor(B1_machine))

## Sample mean intensity dataframe for HCC1806/HS578T diaPASEF 
B2_sample_mean_intensity <- data.frame("mean.intensity" = rowMeans(t(na.omit(B2_data))))
B2_sample_mean_df <- data.frame('var' = row.names(B2_sample_mean_intensity), "mean_intensity" = B2_sample_mean_intensity, "class" = factor(B1_group), "batch" = factor(B1_machine))

### Function for sample mean intensity plot
sample_mean_intensity_plot <- function(mean_intensity_df){
sample_mean_plot <- ggplot(data = mean_intensity_df, aes(x = factor(rownames(mean_intensity_df), levels = rownames(mean_intensity_df)), y = mean.intensity, color = class, shape = batch)) + 
  geom_point(size = 5) + 
  scale_color_manual(values = c("#009E94","#FFA507")) + 
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(legend.position = "none") 

return(sample_mean_plot)
}

## Plot for A549/K562 DDA PASEF 
sample_mean_intensity_plot(A1_sample_mean_df)

## Plot for A549/K562 diaPASEF 
sample_mean_intensity_plot(A2_sample_mean_df)

## Plot for HCC1806/HS578T DDA PASEF 
sample_mean_intensity_plot(B1_sample_mean_df)

## Plot for HCC1806/HS578T diaPASEF 
sample_mean_intensity_plot(B2_sample_mean_df)
```

## PCA plot for batch effects 
```{r}
pcaplot <- function(data, group, batch){
  pc_res <- prcomp(data, scale = TRUE)
  pca_data <- data.frame(Class = group, Batch = batch, X = pc_res$x[,1], Y = pc_res$x[,2])
  pca_data['Batch'] <- as.factor(pca_data$Batch)
  pca1 <- pca_data[pca_data$Batch == 1,]
  pca2 <- pca_data[pca_data$Batch == 2,]
  summary <- summary(pc_res)
  str1 = "PC1:" 
  str2 = "PC2:"
  perc = "%"
  pc1_var = round(summary$importance[2,][1]*100, 1)
  pc2_var = round(summary$importance[2,][2]*100, 1)
  
  plot <- ggplot() + 
  geom_point(data = pca1,aes(fill =Class,x=X, y=Y), pch = 21, size = 8, colour = 'black') +
  geom_point(data = pca2,aes(fill =Class,x=X, y=Y), pch = 24, size = 8, colour = 'black') +
 scale_fill_manual(values = c("#469B94", "#F2A83D")) + 
  theme(axis.title = element_text(size = 18)) +
  theme(axis.text = element_text(size = 21)) +
  theme(legend.title = element_text(size = 15)) +
  theme(plot.title = element_text(size = 18)) + 
  theme(legend.position = "none") + 
  theme(plot.title = element_blank()) + 
  theme(panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlab(paste(str1,pc1_var,perc)) + 
  ylab(paste(str2,pc2_var,perc)) 
  
  return(plot)
  print(summary)
}  

#PCA plot for A549/K562 DDA PASEF 
pca_A1 <- pcaplot(t(A1_data_nonNA), A1_group, A1_machine) #provide na.omit dataframes with class and batch factors 

#PCA plot for A549/K562 diaPASEF 
pca_A2 <- pcaplot(t(A2_data_nonNA), A1_group, A1_machine) #provide na.omit dataframes with class and batch factors 

#PCA plot for HCC1806/HS578T DDA PASEF  
pca_B1 <- pcaplot(t(B1_data_nonNA), B1_group, B1_machine) #provide na.omit dataframes with class and batch factors 

#PCA plot for HCC1806/HS578T diaPASEF  
pca_B2 <- pcaplot(t(B2_data_nonNA), B1_group, B1_machine) #provide na.omit dataframes with class and batch factors 

pca_combined <- ggarrange(pca_A1, pca_A2,pca_B1, pca_B2,
          ncol = 2, nrow = 2)
```

## PVCA plot for batch effects 
```{r}
pvca_plot <- function(data, sample_annotation_meta){ #for this function, use the sample_annotation declared earlier as metadata instead
  pvca_df_res <- prepare_PVCA_df(data, sample_annotation_meta, 
  technical_factors = c("machine"),
  biological_factors = c("class"), 
  pca_threshold = .6, variance_threshold = .01, fill_the_missing = -1)
  colors_for_bars = c('gray1', '#2C4F8B','gray90','#FD7D79')
  names(colors_for_bars) = c('residual', 'biological','biol:techn','technical')
  pvcaplot <- plot_PVCA.df(pvca_df_res, colors_for_bars) + ylim(0,1)
  
return(pvcaplot)
}

#Plot for A549/K562 DDA PASEF 
pvca_plot(A1_data, A1_sampleannotation)

#Plot for A549/K562 diaPASEF 
pvca_plot(A2_data, A1_sampleannotation)

#Plot for HCC1806/HS578T DDA PASEF 
pvca_plot(B1_data, B1_sampleannotation)

#Plot for HCC1806/HS578T diaPASEF 
pvca_plot(B2_data, B1_sampleannotation)
```
